New lower bounds for Schur and weak Schur numbers

### 2021-04-07 article/sat.tex

\par
First

@@

More precisely, we verified the following results using the glucose SAT solver [reference].

>>>>

\begin{notation}
Let \(P\) be a predicate. We denote
\end{notation}

\par
\hyperlink{label4}{\textbf{[4]}} verified with a SAT solver that there are no partition in 5 weakly sum-free subsets of \([\![1,197]\!]\) with
\(A_5 = \{67, 68\} \cup [\![70,134]\!] \cup \{136\}\) in 17 hours and could not provide a similar result when only assuming \(m_5 = 67\) even after several
weeks of runtime. By using the same method as above, we were able to verify that \(WS(5 | m_5 = 67) = 196\) in 0.5 seconds with the SAT solver glucose
on a 2.60 GHz Intel i7 processor PC (we used the non-parallel version here but in the rest of this subsection, we used the parallel version of glucose).
The additional information we gave to the SAT solver is that every partition of \([\![1,66]\!]\) in 4 weakly sum-free subsets starts with a partition of
\([\![1,23]\!]\) in 3 weakly sum-free subsets (this can be checked in a few dozens of minutes with a SAT solver). Among the 3 partitions of \([\![1,23]\!]\) in
3 weakly sum-free subsets, every number always appears in the same subset except for 16 and 17 which can appear in two different subsets. We hardcoded
this external knowledge in the propositional formula which allowed us to gain several orders of magnitude in computation time. We also give the stronger
following result.

@@

More precisely, we verified the following results (\(max~m_5\) is the greatest value of \(m_5\) for which we have not verified that \(WS(5 | m_5) \leq 196\)).

### 2021-04-07 article/schur.tex
commit avant pull

\begin{theorem}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S(n+k) \geq SE_+(k+1)S(n) \]
\end{theorem}

Using a SAT solvers, we are able to exhibit adequate Schur partitions, i.e satisfying \(SE_+\) constraint, 
hence providing lower bound on \(SE_+\). We refer to such partition as \textit{templates}. Moreover, when using 
a template in the above formula, it is possible to get an additive constant. We seek templates providing "affine"
inequalities : \(aS(n+k) + b \) where \(a\) is a lower bound for \(SE_+(k+1)\). Further details can be found in the \hyperref[SAT]{SAT section}.

@@

The previous inequalities gives new lower bounds for \(S(n)\) for
\( n \in [\![9,+\infty]\!] \backslash \{13\} \). We compute the lower
bounds values for \( n \in [\![8,15]\!] \) using the four different inequalities, please notice that the values for \( n = 8\) and \(n = 13\) were already obtained thanks to the first one, found by Rowley. The best lower bounds are highlighted.\\
\\
\begin{center}
\begin{tabular}{|*{5}{c|}}
    \hline
	 & 8 & 9 & 10 & 11 \\
	\hline
	\(33S(n-3) + 6 \) & \cellcolor{yellow} 5286 & 17694 & 55446 & 174444\\
	\hline
	\(111S(n-4) + 43 \) & 4927 & \cellcolor{yellow} 17803 & 59539 & 186523\\
	\hline
	\(380S(n-5) + 148 \) & 5088 & 16868 & \cellcolor{yellow} 60948 & \cellcolor{yellow} 203828 \\
	\hline
	\(1140S(n-6) + 528 \) & 5088 & 15348 & 50688 & 182928\\
	\hline
	\hline
	& 12 & 13 & 14 & 15 \\
	\hline
	\(33S(n-3) + 6 \) & 587505 & \cellcolor{yellow} 2011290 & 6726330 & 21072090\\
	\hline
	\(111S(n-4) + 43 \) & 586789 & 1976176 & 6765271 & 22624951 \\
	\hline
	\(380S(n-5) + 148 \) & \cellcolor{yellow} 638548 & 2008828 & \cellcolor{yellow} 6765288 & \cellcolor{yellow} 23160388 \\
	\hline
	\(1140S(n-6) + 528 \) & 611568 & 1915728 & 6026568 & 20295948 \\
	\hline
\end{tabular}
\end{center}

Except for 8, 9 and 13, the best lower bounds are obtained thanks to the third inequality \( 380S(n-5) + 148 \leqslant S(n) \). The table doesn't go any further, but the same inequality allows to improve the lower bounds for every \( n \in [\![15,+\infty]\!] \).

>>>>

\begin{theorem}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S(n+k) \geqslant SE_+(k+1)S(n) \]
\end{theorem}

Using a SAT solvers, we are able to exhibit adequate Schur partitions, i.e satisfying \(SE_+\) constraint, 
hence providing lower bound on \(SE_+\). We refer to such partition as \textit{templates}. Moreover, when using 
a template in the above formula, it is possible to get an additive constant. We seek templates providing "affine"
inequalities : \(aS(n+k) + b \) where \(a\) is a lower bound for \(SE_+(k+1)\). Further details can be found in the \hyperref[SAT]{SAT section}.
Here are the best inequalities on Schur numbers so far :
\[ S(n+1) \geq 3S(n) + 1 \\
 S(n+2) \geq 9S(n) + 4 \\
 S(n+3) \geq 33S(n) + 6 \\
 S(n+4) \geq 111S(n) + 43 \\
 S(n+5) \geq 380S(n) + 148 \\
 S(n+6) \geq 1140S(n) + 528 \]

@@

The previous inequalities gives new lower bounds for \(S(n)\) for
\( n \in [\![9,+\infty]\!] \backslash \{13\} \). We compute the lower
bounds values for \( n \in [\![8,15]\!] \) using the four different inequalities, please notice that the values for \( n = 8\) and \(n = 13\) were already obtained thanks to the first one, found by Rowley. The best lower bounds are highlighted.\\
\\
\begin{center}
\begin{tabular}{|*{5}{c|}}
    \hline
	 & 8 & 9 & 10 & 11 \\
	\hline
	\(33S(n-3) + 6 \) & \cellcolor{yellow} 5286 & 17694 & 55446 & 174444\\
	\hline
	\(111S(n-4) + 43 \) & 4927 & \cellcolor{yellow} 17803 & 59539 & 186523\\
	\hline
	\(380S(n-5) + 148 \) & 5088 & 16868 & \cellcolor{yellow} 60948 & \cellcolor{yellow} 203828 \\
	\hline
	\(1140S(n-6) + 528 \) & 5088 & 15348 & 50688 & 182928\\
	\hline
	\hline
	& 12 & 13 & 14 & 15 \\
	\hline
	\(33S(n-3) + 6 \) & 587505 & \cellcolor{yellow} 2011290 & 6726330 & 21072090\\
	\hline
	\(111S(n-4) + 43 \) & 586789 & 1976176 & 6765271 & 22624951 \\
	\hline
	\(380S(n-5) + 148 \) & \cellcolor{yellow} 638548 & 2008828 & \cellcolor{yellow} 6765288 & \cellcolor{yellow} 23160388 \\
	\hline
	\(1140S(n-6) + 528 \) & 611568 & 1915728 & 6026568 & 20295948 \\
	\hline
\end{tabular}
\end{center}
Except for 8, 9 and 13, the best lower bounds are obtained thanks to the third inequality \( 380S(n-5) + 148 \leqslant S(n) \). The table doesn't go any further, but the same inequality allows to improve the lower bounds for every \( n \in [\![15,+\infty]\!] \).
